---
title: "Hospital-Readmission-Model"
author: "Lianna Novitz"
date: "5/1/2018"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(readr)
library(gridExtra)
library(GGally) 
library(leaps) 
library(glmnet)
library(faraway)
library(forcats) #to collapse categorical levels
library(ISLR)
library(rpart)
library(rpart.plot)
library(randomForest)
library(knitr)
```

```{r}
diabetes <- read_csv("https://www.dropbox.com/s/odtsp7foaugu694/diabetes_data_clean.csv?dl=1")
```

###Introduction
According to the authors, “The dataset represents 10 years (1999-2008) of clinical care at 130 US hospitals and integrated delivery networks. It includes over 50 features representing patient and hospital outcomes. Information was extracted from the database for encounters that satisfied the following criteria.

1. It is an inpatient encounter (a hospital admission).
2. It is a diabetic encounter, that is, one during which any kind of diabetes was entered to the system as a diagnosis.
3. The length of stay was at least 1 day and at most 14 days.
4. Laboratory tests were performed during the encounter.
5. Medications were administered during the encounter."

Citation: Beata Strack, Jonathan P. DeShazo, Chris Gennings, Juan L. Olmo, Sebastian Ventura, Krzysztof J. Cios, and John N. Clore, “Impact of HbA1c Measurement on Hospital Readmission Rates: Analysis of 70,000 Clinical Database Patient Records,” BioMed Research International, vol. 2014, Article ID 781670, 11 pages, 2014

###Research Question
It's a problem for hospitals if a patient is readmitted before 30 days, so my question is, what characteristics make a patient more likely to be readmitted before 30 days? How can hospital staff reduce this rate?

First of all, let's see what the rate of readmission before 30 days is.

```{r}
kable(
  diabetes %>% 
    group_by(readmitted) %>% 
    count(),
  align = c("l", "c"),
  caption = "Table 1: Readmission Categories"
)
```
```{r, echo=FALSE}
11357/(11357 + 35545 + 54864)
35545/(11357 + 35545 + 54864)
54864/(11357 + 35545 + 54864)
```
The rate of readmission before 30 days is about 11%. The rate of readmission after 30 days is about 35%. The rate of no readmission is about 54%.

So our job is to find the features which make a hospital patient more likely to be readmitted before 30 days. 

###Exploring Relationships


```{r, warning=FALSE, message=FALSE}
summary(diabetes)
```


```{r, echo=FALSE}
diabetes %>% 
    group_by(race) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(gender) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(age) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(time_in_hospital) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(num_lab_procedures) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(num_procedures) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(num_medications) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_outpatient) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_emergency) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_inpatient) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_diagnoses) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(max_glu_serum) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(A1Cresult) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(metformin) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(glimepiride) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(glipizide) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(glyburide) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(pioglitazone) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(rosiglitazone) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(insulin) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(change) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(diabetesMed) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(readmitted) %>% 
    count()
```
```{r}
11357 + 35545
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(admission_source) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(admission_type) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(discharge_disposition) %>% 
    count()
```
I want to find out characteristics about patients who were readmitted before 30 days was up. 

The density plot below may hint that patients who were sent home have slightly higher rates of readmittance before 30 days. 
```{r}
diabetes %>% 
  ggplot(aes(x=discharge_disposition, fill=factor(readmitted))) +
  geom_density(alpha=.5)
```

The boxplot below shows, perhaps, a slightly higher rate of readmission before 30 days is correlated with a higher number of inpatient visits in the year preceding the specific hospital this specific encounter. This intuitively makes sense.

```{r}
diabetes %>% 
  ggplot(aes(x=factor(readmitted), y=number_inpatient)) +
  geom_boxplot()
```
The boxplot below does seem to show a difference between the amount of time spent in the hospital for those readmitted after the encounter and those not readmitted.
```{r}
diabetes %>% 
  ggplot(aes(x=factor(readmitted), y=time_in_hospital)) +
  geom_boxplot()
```

### Decision Tree Model
```{r}
earlyadmit_tree <- rpart(readmitted ~ .,
                 data=diabetes)
```

```{r}
rpart.plot(earlyadmit_tree, under=FALSE, varlen = 0, yesno = 2, ycompress = TRUE)
```

Another tree to see if I can get better detail in the visuals
```{r}
rpart.plot(earlyadmit_tree, # middle graph
type=4,
extra=101, 
box.palette="GnBu",
branch.lty=3, 
shadow.col="gray", 
nn=TRUE
)
```

```{r}
earlyadmit2_tree <- rpart(readmitted ~ discharge_disposition,
                 data=diabetes)
```

```{r}
rpart.plot(earlyadmit2_tree)
```

### Logistic Model

Before fitting a logistic model, I transformed the readmitted variable so it only has two levels. 

Before the transformation, the readmitted variable contained 3 levels, revealing whether a patient was readmitted before 30 days, after 30 days, or not at all after the recorded encounter.

```{r}
diabetes %>% 
    group_by(readmitted) %>% 
    count()
```

After the transformation, the readmitted variable contains only 2 levels, revealing whether a patient was readmitted at all or not at all after the recorded encounter.
```{r, echo=FALSE}
diabetes2 <- diabetes %>% 
  mutate(readmitted = readmitted %>% fct_collapse(YES = c("<30",">30")))
```

We also decided to transform the readmitted variable in a different way that contains only 2 levels. Instead of collapsing readmissions into one level, I will treat readmissions after thirty days and no readmissions as one level, and readmissions before thirty days as the second level.
```{r, echo=FALSE}
diabetes3 <- diabetes %>% 
  mutate(readmitted = readmitted %>% fct_collapse(NO = c("NO",">30")))
```

```{r}
diabetes2 %>% 
    group_by(readmitted) %>% 
    count()
```

```{r}
log.admit <- glm(readmitted ~ .,
               data=diabetes2, family = binomial(link = "logit"))
```

```{r}
summary(log.admit)
```

What is a nonparametric model?
