---
title: "Hospital-Readmission-Model"
author: "Lianna Novitz"
date: "5/1/2018"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(readr)
library(gridExtra)
library(GGally) 
library(leaps) 
library(glmnet)
library(faraway)
library(forcats) #to collapse categorical levels
library(ISLR)
library(knitr)
```

```{r}
diabetes <- read_csv("https://www.dropbox.com/s/odtsp7foaugu694/diabetes_data_clean.csv?dl=1")
```

###Introduction
According to the authors, “The dataset represents 10 years (1999-2008) of clinical care at 130 US hospitals and integrated delivery networks. It includes over 50 features representing patient and hospital outcomes. Information was extracted from the database for encounters that satisfied the following criteria.

1. It is an inpatient encounter (a hospital admission).
2. It is a diabetic encounter, that is, one during which any kind of diabetes was entered to the system as a diagnosis.
3. The length of stay was at least 1 day and at most 14 days.
4. Laboratory tests were performed during the encounter.
5. Medications were administered during the encounter."

Citation: Beata Strack, Jonathan P. DeShazo, Chris Gennings, Juan L. Olmo, Sebastian Ventura, Krzysztof J. Cios, and John N. Clore, “Impact of HbA1c Measurement on Hospital Readmission Rates: Analysis of 70,000 Clinical Database Patient Records,” BioMed Research International, vol. 2014, Article ID 781670, 11 pages, 2014

###Research Question
It's a problem for hospitals if a patient is readmitted before 30 days, so my question is, what characteristics make a patient more likely to be readmitted before 30 days? How can hospital staff reduce this rate?

First of all, let's see what the rate of readmission before 30 days is.

```{r}
kable(
  diabetes %>% 
    group_by(readmitted) %>% 
    count(),
  align = c("l", "c"),
  caption = "Table 1: Readmission Categories"
)
```
```{r, echo=FALSE}
11357/(11357 + 35545 + 54864)
35545/(11357 + 35545 + 54864)
54864/(11357 + 35545 + 54864)
```
The rate of readmission before 30 days is about 11%. The rate of readmission after 30 days is about 35%. The rate of no readmission is about 54%.

So our job is to find the features which make a hospital patient more likely to be readmitted before 30 days. 

###Exploring Relationships


```{r, warning=FALSE, message=FALSE}
summary(diabetes)
```


```{r, echo=FALSE}
diabetes %>% 
    group_by(race) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(gender) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(age) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(time_in_hospital) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(num_lab_procedures) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(num_procedures) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(num_medications) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_outpatient) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_emergency) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_inpatient) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(number_diagnoses) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(max_glu_serum) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(A1Cresult) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(metformin) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(glimepiride) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(glipizide) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(glyburide) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(pioglitazone) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(rosiglitazone) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(insulin) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(change) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(diabetesMed) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(readmitted) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(admission_source) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(admission_type) %>% 
    count()
```

```{r, echo=FALSE}
diabetes %>% 
    group_by(discharge_disposition) %>% 
    count()
```
We want to find out characteristics about patients who were readmitted before 30 days was up. 

The boxplot below shows, perhaps, a slightly higher rate of readmission before 30 days is correlated with a higher number of inpatient visits in the year preceding the specific hospital this specific encounter. This intuitively makes sense.

```{r}
diabetes %>% 
  ggplot(aes(x=factor(readmitted), y=number_inpatient)) +
  geom_boxplot()
```
The boxplot below does not seem to show a difference between the amount of time spent in the hospital for those readmitted within thirty days of the encounter versus those readmitted after thirty days of the encounter. There does appear to be a difference between the amount of time spent in the hospital for those not readmitted after the encounter in that year and those readmitted, however.
```{r}
diabetes %>% 
  ggplot(aes(x=factor(readmitted), y=time_in_hospital)) +
  geom_boxplot()
```

### Analysis

Using a logistic model, our goal is to predict readmittance before thirty days. We are confined to a binary response variable, so we have to combine levels within readmitted to build a proper model. 

We transformed the readmitted variable so it only has two levels. 

Before the transformation, the readmitted variable contained 3 levels, revealing whether a patient was readmitted before 30 days, after 30 days, or not at all after the recorded encounter.

```{r}
diabetes %>% 
    group_by(readmitted) %>% 
    count()
```

After the transformation, the readmitted variable contains only 2 levels, revealing whether a patient was readmitted at all or not at all after the recorded encounter. We'll call this version 1 of the readmitted variable.
```{r, echo=FALSE}
diabetes2 <- diabetes %>% 
  mutate(readmitted = readmitted %>% fct_collapse(YES = c("<30",">30")))
```

```{r}
diabetes2 %>% 
    group_by(readmitted) %>% 
  count()
```

We also decided to transform the readmitted variable in a different way that contains only 2 levels. 

Instead of collapsing readmissions into one level, I will treat readmissions after thirty days and no readmissions as one level, and readmissions before thirty days as the second level. We'll call this version 2 of the readmitted variable.
```{r, echo=FALSE}
diabetes3 <- diabetes %>% 
  mutate(readmitted = readmitted %>% fct_collapse(NO = c("NO",">30")))
```

```{r}
diabetes3 %>% 
    group_by(readmitted) %>% 
    count()
```

Now to build our logistic model, we will use the lasso method to control for how many variables we include in the model. First we divide the data into a training and validation set. 

The code below uses the dataset with version 1 of the readmitted variable.
```{r}
diabetes.trval <- diabetes2 %>% 
  mutate(grp = sample(c("train", "validation"), 
                      size=n(), 
                      prob = c(.6,.4),
                      replace=TRUE))
train <- diabetes.trval %>% filter(grp=="train")
validation <- diabetes.trval %>% filter(grp=="validation")
```


The code below uses the dataset with version 2 of the readmitted variable.
```{r}
diabetes.trval2 <- diabetes3 %>% 
  mutate(grp = sample(c("train", "validation"), 
                      size=n(), 
                      prob = c(.6,.4),
                      replace=TRUE))
train2 <- diabetes.trval2 %>% filter(grp=="train")
validation2 <- diabetes.trval2 %>% filter(grp=="validation")
```

Then we use the lasso method to build our two logistic models representing two different versions of the readmitted variable.
```{r}
x2 <- model.matrix(readmitted ~., data=train %>% select(-grp))[,-1]
y2 <- train$readmitted

x3 <- model.matrix(readmitted ~., data=train2 %>% select(-grp))[,-1]
y3 <- train2$readmitted

set.seed(1)
cv.lasso.mod <- cv.glmnet(x = x2, y = y2, alpha = 1, family="binomial")
cv.lasso.mod2 <- cv.glmnet(x = x3, y = y3, alpha = 1, family="binomial")
```

```{r, echo=FALSE}
lasso.stats <- tibble(lambda = cv.lasso.mod$lambda,
                      cverror = cv.lasso.mod$cvm,
                      upper = cv.lasso.mod$cvup,
                      lower = cv.lasso.mod$cvlo) 

lasso.stats2 <- tibble(lambda = cv.lasso.mod2$lambda,
                      cverror = cv.lasso.mod2$cvm,
                      upper = cv.lasso.mod2$cvup,
                      lower = cv.lasso.mod2$cvlo) 
```

Below is a plot showing optimal lambda values that minimize the test error rate for our first model. As the lambda value increases, the test error rate changes more rapidly than what we see in our second model (see below.)
```{r}
lasso.stats %>% 
  ggplot(aes(x=lambda, y=cverror)) +
  geom_point(color="red") +
  geom_errorbar(aes(ymin=lower, ymax=upper)) +
  xlim(0, 0.008) +
  geom_vline(xintercept = cv.lasso.mod$lambda.1se, color="orange", alpha=.5)
```
Below is a plot showing optimal lambda values that minimize the test error rate for our second model. As we can see, as the lambda value increases, the test error rate remains pretty constant, which is good if we want to make our model more simple with less variables.

```{r}
lasso.stats2 %>% 
  ggplot(aes(x=lambda, y=cverror)) +
  geom_point(color="blue") +
  geom_errorbar(aes(ymin=lower, ymax=upper)) +
  xlim(0, 0.008) +
  geom_vline(xintercept = cv.lasso.mod2$lambda.1se, color="orange", alpha=.5)
```

To see the models that were fit, we use the *predict* function with three different values of *lambda*. Below, we look at the models for the value of $\lambda$ with the smallest test error, the value of $\lambda$ with the largest test error within one standard deviation of the smallest test error, and $\lambda= 0.0004$, which is between the two. 

```{r}
predict(cv.lasso.mod, 
        type="coefficients", 
        s = c(cv.lasso.mod$lambda.min, cv.lasso.mod$lambda.1se, 0.008))
```

```{r}
predict(cv.lasso.mod2, 
        type="coefficients", 
        s = c(cv.lasso.mod2$lambda.min, cv.lasso.mod2$lambda.1se, 0.0075))
```
Using a lambda value of 0.0075, the lasso method gives us a model with only 4 variables, which is quite simple. This model is for Version 2 of the readmitted variable, which treats patients readmitted before 30 days as its own level, and all other patients (readmitted after 30 days or not readmitted at all) as a separate level. 

Below we use the first version of the readmitted variable to calculate the test error rate of our model with an optimized value of lambda. 

Model 1's Test Error Rate
```{r}
newX <- model.matrix(factor(readmitted) ~., data = validation[-c(27)])[,-1] #diabetes3
pred = predict(cv.lasso.mod,newx = newX, type = "response")
pred_class <- as.numeric(pred > 0.5)
table(validation$readmitted, pred_class)
error = (10375 + 4607)/(8302 + 10375 + 4607 + 17336)
error
```

Model 2's Test Error Rate
```{r}
newX <- model.matrix(factor(readmitted) ~., data = validation2[-c(27)])[,-1] #diabetes3
pred = predict(cv.lasso.mod2,newx = newX, type = "response", s = c(0.0075))
pred_class <- as.numeric(pred > 0.5)
table(validation2$readmitted, pred_class)
error = (4483 + 54)/(47 + 4483 + 54 + 35983)
error
```

Since the test error rate of Model 1 is nearly 37%, we're going to choose Model 2 as the best predictive model for readmission rates because Model 2's test error is only about 11%. Model 2's equation is below.
$$
readmitted = -0.00066(numberemergency) + -0.24(numberinpatient) + \\ -0.0064(numberdiagnoses)
+ -0.37*I(dischargedisposition)
$$

Note: I is 1 if the patient was transferred to another facility following the incident.

Next we compute the odds that a diabetes patient is readmitted to the hospital within 30 days based on the number of emergency visits recorded prior to the incident, the number of inpatient visits recorded prior to the incident, the number of diagnoses made at the hospital, and whether the patient was transferred to another hospital following the incident.

```{r}
odds_emer = exp(coef(cv.lasso.mod2, s = c(0.0075))[23])
odds_emer
```
The odds of a patient being readmitted to the hospital within 30 days are multiplied by a factor of .99, given the number of emergency room visits.
```{r}
odds_inpat = exp(coef(cv.lasso.mod2, s = c(0.0075))[24])
odds_inpat
```
The odds of a patient being readmitted to the hospital within 30 days are multiplied by a factor of .79, given the number of inpatient visits.
```{r}
odds_diag = exp(coef(cv.lasso.mod2, s = c(0.0075))[25])
odds_diag
```
The odds of a patient being readmitted to the hospital within 30 days are multiplied by a factor of .99, given the number of diagnoses.
```{r}
odds_transfer = exp(coef(cv.lasso.mod2, s = c(0.0075))[64])
odds_transfer
```
The odds of a patient being readmitted to the hospital within 30 days are multiplied by a factor of .69, given that the patient was transferred to another facility.

### Conclusion

In conclusion, we built a logistic model to help predict hospital readmission rates for diabetes patients. We defined readmission with two variable transformations. After using the lasso method to compare model error rates, we ended up choosing the definition that either a patient was readmitted within 30 days or the patient was not. Our final model is the following: $readmitted = -0.00066(numberemergency) + -0.24(numberinpatient) + -0.0064(numberdiagnoses)+ -0.37*I(dischargedisposition)$. The number of emergency room and inpatient visits prior to the incident, number of diagnoses, and whether the patient was transferred to another facility ended up being important variables to predict hospital readmission.

### Appendix

```{r, eval=FALSE}
diabetes2 <- diabetes %>% 
  mutate(readmitted = readmitted %>% fct_collapse(YES = c("<30",">30")))
```

```{r, eval=FALSE}
diabetes3 <- diabetes %>% 
  mutate(readmitted = readmitted %>% fct_collapse(NO = c("NO",">30")))
```